---
title: "Hooper_ESM244_HW2_Task3"
author: "Tess Hooper"
date: "2/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

## Load packages

library(tidyverse)
library(sf)
library(tmap)
library(ggspatial)
library(janitor)
library(here)
library(nationalparkcolors)
library(paletteer)
library(palettetown)
library(ggsci)
library(leaflet)
library(shinyjs)
library(gameofthrones)


## Read in data 

lu_lc <- read_sf(dsn = here("Land_Use_Land_Cover_LULC"), layer = "Land_Use_Land_Cover_LULC") %>% 
  select(landcover, st_areasha, perimeter) %>% # Sticky geometry so don't need to select that
  st_transform(crs = 4326) # Ensure in right CRS

watersheds <- read_sf(dsn = here("Watersheds"), layer = "Watersheds") %>% 
  select(wuname, area_sqmi, hucarea, st_areasha, st_perimet) %>% # Sticky geometry so don't need to select that
  st_transform(crs = 4326)

```


### **Land Use/Land Cover and Watersheds on the Big Island, Hawaii**


**Citations:**  
*Land use/land cover data:* http://geoportal.hawaii.gov/datasets/land-use-land-cover-lulc
*Watershed data:* http://geoportal.hawaii.gov/datasets/watersheds

#### 1. Land Use/Land Cover 
```{r}

### Wrangling and Exploration:

# Look at LULC: 
plot(lu_lc)


#unique(lu_lc$landcover) 
# 28 different land use types. One of them is just "0" -- does that mean no data available? 
# Could group land use types into categories: Agriculture, Residential, Urban, Aquatic?

#### TYPES ####
#[1] "Cropland and Pasture"                                                     
#[2] "Commercial and Services"                                                  
#[3] "Residential"                                                              
#[4] "Evergreen Forest Land"                                                    
#[5] "Other Urban or Built-up Land"                                             
#[6] "Mixed Rangeland"                                                          
#[7] "Industrial"                                                               
#[8] "Streams and Canals"                                                       
#[9] "Orchards, Groves, Vineyards, Nurseries and Ornamental Horticultural Areas"
#[10] "Shrub and Brush Rangeland"                                                
#[11] "Forested Wetland"                                                         
#[12] "Reservoirs"                                                               
#[13] "Nonforested Wetland"                                                      
#[14] "Bare Exposed Rock"                                                        
#[15] "Sandy Areas Other than Beaches"                                           
#[16] "Transportation, Communications and Utilities"                             
#[17] "Herbaceous Rangeland"                                                     
#[18] "Beaches"                                                                  
#[19] "Other Agricultural Land"                                                  
#[20] "Lakes"                                                                    
#[21] "Strip Mines, Quarries, and Gravel Pits"                                   
#[22] "Mixed Barren Land"                                                        
#[23] "Bays and Estuaries"                                                       
#[24] "Mixed Urban or Built-up Land"                                             
#[25] "Transitional Areas"                                                       
#[26] "0"                                                                        
#[27] "Industrial and Commercial Complexes"                                      
#[28] "Confined Feeding Operations"  

lu_lc_categories <- lu_lc %>% 
  mutate(
    landcover = case_when (
      landcover %in% c("Industrial and Commercial Complexes",
                        "Mixed Urban or Built-up Land",
                        "Transportation, Communications and Utilities",
                        "Industrial",
                        "Commercial and Services",
                        "Other Urban or Built-up Land" ) ~ "Urban", #Group this into a broader "Urban" category 
      landcover %in% c("Confined Feeding Operations",
                        "Other Agricultural Land",
                        "Orchards, Groves, Vineyards, Nurseries and Ornamental Horticultural Areas",
                        "Cropland and Pasture") ~ "Agricultural", # Group these into a broader "Agricultral" category
      landcover %in% c("Bays and Estuaries",
                        "Lakes",
                        "Reservoirs",
                        "Streams and Canals") ~ "Waterways", # Group these into a broader "Waterways" category
      landcover %in% c("Nonforested Wetland",
                       "Forested Wetland") ~ "Wetlands", # "Wetlands" category
      landcover %in% c("Mixed Rangeland",
                       "Shrub and Brush Rangeland",
                       "Herbaceous Rangeland") ~ "Rangeland",
      TRUE ~ landcover)) %>%  # Keep everything else as is 
  filter(landcover != 0) # Filter out the "0" landcover (no data?)
  
unique(lu_lc_categories$landcover)
# Now 13 categories, little bit better. Could group more if necessary? 

plot(lu_lc_categories)
# Still have area and perimeter. Not necessary but could be useful in visualizing the total area of each land use type. 

### Maybe only visualize certain land uses??
                        
```

<br>

```{r}

### Visualize Land Use / Land Cover for just Big Island

lulc_plot <- ggplot(data = lu_lc_categories) +
  geom_sf(data = lu_lc_categories, 
          color = NA, # Gets rid of gray coloring borders which made categories difficult to see 
          aes(fill = landcover)) +
  scale_fill_paletteer_d("ggsci::springfield_simpsons") +
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Land Use and Land Cover on Big Island, Hawaii") +
  labs(fill = "Landcover") +
  theme_minimal() +
  coord_sf(xlim = c(-156.5, -154.5), ylim = c(18.5, 20.5), expand = FALSE) + # Clip to just Big Island
  theme(panel.background = element_rect(fill = "paleturquoise")) # Add background light blue color to make it look like ocean. Is it too bright? Need another color maybe. 
  

## Could add more details in this plot? Maybe add ocean color background. Could also zoom in to one island? Is it too difficult to see individual land uses? Or maybe need to decrease categories. 
  

lulc_plot


```
  
**Figure 1. Land Use and Land Cover on Big Island, Hawaii**. Land use and land cover for thirteen different categories. 
<br>

Make it interactive!
```{r}

## Make an interactive map using Tmap


tmap_mode("view") # Tmap set to interactive viewing

lulc_tmap <- tm_basemap("Esri.WorldImagery") + # Add general basemap
  tm_shape(lu_lc_categories) +
  tm_fill(col = "landcover", palette = "Set1") + # Fill by landuse 
  tm_legend(legend.position = c("right", "bottom")) +
  tm_layout(title = "Land Use and Land Cover in Hawaii")

#leaflet::providers$Stamen.Watercolor

lulc_tmap

# World Imagery might not be the best basemap to use. It's a little dark? 



```
  
**Figure 2. Interactive Map for Land Use and Land Cover on Hawaii.**

<br>


#### 2. Watersheds in Hawaii
```{r}

## Plot watersheds

# plot(watersheds)

### Visualize watersheds by area

watersheds_plot <- ggplot(data = watersheds) +
  geom_sf(data = watersheds, 
          color = NA, # Gets rid of gray coloring borders which made categories difficult to see 
          aes(fill = area_sqmi)) +
  scale_fill_paletteer_c("gameofthrones::daenerys") +
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Watersheds on Hawaii") +
  labs(fill = "Area (Sq Mi)") +
  theme(panel.grid.major = element_line(color = gray(.1), 
                                        linetype = "dashed", size = 0.1), 
        panel.background = element_rect(fill = "aliceblue"))

watersheds_plot

```
  
**Figure 3. Watersheds in Hawaii by Area.** Watersheds on Hawaii delineated by area. 

Crop to just Big Island, Hawaii
```{r}


ws_bigisland_plot <- ggplot(data = watersheds) +
  geom_sf(data = watersheds, 
          color = NA, # Gets rid of gray coloring borders which made categories difficult to see 
          aes(fill = area_sqmi)) +
  scale_fill_paletteer_c("gameofthrones::daenerys") +
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Watersheds on Hawaii") +
  labs(fill = "Area (Sq Mi)") +
  theme(panel.grid.major = element_line(color = gray(.1), 
                                        linetype = "dashed", size = 0.1), 
        panel.background = element_rect(fill = "aliceblue")) +
  coord_sf(xlim = c(-156.5, -154.5), ylim = c(18.5, 20.5), expand = FALSE)

ws_bigisland_plot



```
  
**Figure 4. Watersheds on Big Island, Hawaii.** Watersheds on Big Island, Hawaii delineated by area. 
